### 2020-12-28 paper/paper.tex
fix: some writing fixes

\documentclass[english, 12pt]{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{dsfont}
\usepackage{enumitem}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage[onehalfspacing]{setspace}

\theoremstyle{definition}
\newtheorem{example}{Example}[section]

\theoremstyle{definition}
\newtheorem{definition}{Definition}

\theoremstyle{definition}
\newtheorem{theorem}{Theorem}

\begin{document}

% TODO title
\title{A bad day of studying is better than a good day of sleeping}
\subtitle{Report on 'How to Revise a Total Preorder' \cite{Booth2011} for seminar 1901 - Representation and processing of uncertain knowledge with logic-based methods}
\author{
	Heltweg, Philip (3230880) \\
	University of Hagen \\
	pheltweg@gmail.com
}
\date{\today}
\maketitle

% TODO abstract
\abstract{Abstract goes here}

\newpage

\tableofcontents

\newpage

% TODO introduction
\section{Introduction}

% goal of paper: interested in function * for getting new tpo

% roten faden des reports beschreiben

\subsection{Research context}
A core process for an agent dealing with uncertain knowledge is updating their worldview when new information becomes available, also called belief change. As an agent can be human or a machine, work in belief change has impact on both philosophy and artificial intelligence \cite{Ferme2011}. Different approaches to handling belief change have been discussed with notable approaches including nonmonotonic logic, probabilistic reasoning and belief revision \cite{Darwiche1997}.

In belief revision, belief change is modelled using an operator that produces an updated set of beliefs from the state of an agent and new evidence. Research in this area discusses appropriate formalisations of agent state and constrains on individual operators or families of operators. Depending on if the suggested operator handles only one new evidence or multiple new evidences the approaches are called one-step or iterated belief revision.

%Consider a knowledge base (KB) represented by a set of sentences in a
%language L. As our perception of the world described by the knowledge base
%changes, the knowledge base must be modified. G~irdenfors [9] distinguishes
%several kinds of modifications. If we simply acquire additional knowledge
%about the world, and the new knowledge does not conflict with the current
%beliefs t of the KB, we expand the KB. If, however, the new knowledge is
%inconsistent with the old beliefs, and we want the KB to be always consistent,
%we must resolve the conflict somehow; this operation will be called revision. A
%* Part of this work was performed while this author was visiting the University of Toronto.
%** This author is partially supported by the Natural Science and Engineering Research Council
%of Canada.
%1 We use the terms belief and knowledge interchangeably in this paper.
%0004-3702/91/$03.50 © 1991 -- Elsevier Science Publishers B.V. All rights reserved 
%264 H. Katsuno, A.O. Mendelzon
%different kind of change occurs when a sentence previously believed becomes
%questionable and has to be given up; the operation that makes this change is a
%contraction

% TODO: discussed in \cite{Katsuno1991} correct? was mentioned in intro of Admissible and Restrained Revision
A common tool used to encode plausibility assumptions of an agent about possible worlds are total preorders (shortened as tpos) \cite{Booth2011}, discussed in \cite{Katsuno1991}. The primary article for this report, 'How to Revise a Total Preorder', researches how to model the change of these total preorders when new evidence becomes available. Being able to derive a new tpo from additional evidence is an approach to handling multiple revision steps and therefor places the article in the field of iterated belief revision.
%1.2 Ich sehe hier die Gefahr, dass das zu Ausführlich wird.
%Keine große Diskussion von AGM, Darwiche and Pearl und so weiter.
%Keine Beziehung zur Allen und co herstellen ;)
%Kurz und Prägnant präsentieren, was benötgigt wird um zu Verstehen was
%in dem Primärartikel neu ist.

\section{Formal background}

\subsection{Propositional Logic} % Aussagenlogik
Notation follows as used in \cite{Booth2011}: A propositional language $L$ generated from finitely many propositional variables, lower case greek letters represent formulae in $L$ and $\top$, $\bot$ representing tautology and contradiction respectively. $\vdash$ as classical logical consequence, $\equiv$ classical logical equivalence. $Cn$ is used to denote the deductive closure of a formula or set of formulae in $L$.

$W$ is the set of propositional worlds (also called propositional interpretations in classical logic \cite{Kai2020}). Given a $\alpha \in L$ the set of worlds that satisfy $\alpha$ is denoted by $[\alpha]$.

For any set of worlds $S \subseteq W$ of worlds, $Th(S)$ is the set of sentences true in all worlds in $S$.
% TODO Th(S) not needed?

\subsection{Total preorders}
A total preorder is a binary relation $\leq$ (here over the set of propositional worlds $W$) that is total, reflexive and transitive (i.e. for all $x, y, z \in W$ either $x \leq y$ or $y \leq x$, $x \leq x$ and if $x \leq y$ and $y \leq z$ then $x \leq z$).

The symbol $<$ is used to denote the strict part of $\leq$ while $\sim$ represents the symmetric closure of $\leq$ (i.e. $x \sim y$ iff $x \leq y$ and $y \leq x$).

Tpos are a common tool to handle preference orderings over propositional worlds in literature about belief change \cite{Booth2011}.

A helpful visualisation for tpos is described in \cite{Booth2006}. It uses the fact that tpos can be represented as a linearly ordered set of ranks. Each rank of a tpo $\leq$ is defined as the equivalence classes modulo the symmetric closure of $\leq$: $[[x]]_{\sim} = \{y \mid y \sim x\}$. These equivalence classes are then ordered by the relation $[[x]] \leq [[y]]$ iff $x \leq y$

\begin{example}
    \label{example:example-introduction}
    As an accompanying example consider the following situation, closely aligned to \cite{Booth2011} and \cite{Darwiche1997}:

    Our agent is a judge in a murder trial. John and Mary are suspects. $p$ represents "John is the murderer", $q$ represents "Mary is the murderer" and $r$ represents "the victim is an alien". Possible propositional worlds will be denoted as triplets of 0s and 1s denoting $p$, $q$ and $r$ to be true or false, e.g. the $101$-world stands for John being the murderer and the victim being an alien.

    To start the agent believes it is reasonable to assume the murderer acted alone (but is not ruling out both of the accused conspiring). In addition it would be extremely surprising (but not impossible) if the victim was an alien.

    Consider the propositional worlds $W = \{ 000, 001, 010, 011, 100, 101, 110, 111\}$. A tpo representing the judges assumptions would be $\leq$ with $010 \sim 100 < 000 \sim 110 < 011 \sim 101 < 001 \sim 111$. The equivalent representation as a linearly ordered set of ranks is shown in table \ref{tab:visualising-a-tpo-example}.

    \begin{table}[h]
         \centering
        \begin{tabular}{llll}
        $R_{1}$                      & $R_{2}$                                                                   & $R_{3}$ & $R_{4}$                      \\ \hline
        \multicolumn{1}{|l|}{\begin{tabular}[c]{@{}l@{}}$010$\\ $100$\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}$000$\\ $110$\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}$011$\\ $101$\end{tabular}} &
            \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}$001$\\ $111$\end{tabular}} \\ \hline
        \end{tabular}
        \caption{Visualising a tpo as a linearly ordered set of ranks as done in \cite{Booth2006}}
        \label{tab:visualising-a-tpo-example}
    \end{table}
\end{example}

\subsection{Belief sets and epistemic states}
A belief set is the deductively closed set of propositions an agent accepts as true at any given point in time \cite{Ferme2011}. AGM theory \cite{Alchourron1985} defines postulates for belief sets and their contraction, revision or expansion with new evidence.
% TODO explain contraction, revision, expansion?

%    As our perception of the world described by the knowledge base
%    changes, the knowledge base must be modified. G~irdenfors [9] distinguishes
%    several kinds of modifications. If we simply acquire additional knowledge
%    about the world, and the new knowledge does not conflict with the current
%    beliefs t of the KB, we expand the KB. If, however, the new knowledge is
%    inconsistent with the old beliefs, and we want the KB to be always consistent,
%    we must resolve the conflict somehow; this operation will be called revision. A
%    * Part of this work was performed while this author was visiting the University of Toronto.
%    ** This author is partially supported by the Natural Science and Engineering Research Council
%    of Canada.
%    1 We use the terms belief and knowledge interchangeably in this paper.
%    0004-3702/91/$03.50 © 1991 -- Elsevier Science Publishers B.V. All rights reserved 
%    264 H. Katsuno, A.O. Mendelzon
%    different kind of change occurs when a sentence previously believed becomes
%    questionable and has to be given up; the operation that makes this change is a
%    contraction. Other kinds of operations are discussed in [14].

Darwiche and Pearl \cite{Darwiche1997} argue that working with belief sets is not expressive enough for satisfying results in iterated belief revision. They make the distinction between propositional beliefs (i.e. beliefs the agent accepts and are part of the belief set) and conditional beliefs (i.e. beliefs the agent is prepared to adopt with new evidence).

While the AGM postulates define restrictions on revising propositional beliefs (a core one being the principle of minimal change), they place little restrictions on preserving conditional beliefs. In their paper Darwiche and Pearl use epistemic states, abstract entities that contain all information an agent needs for it's reasoning, especially their strategy for belief revision. As conditional beliefs are represented by these strategies, it is necessary to define how to modify the strategy itself when encountering multiple new information \cite{Darwiche1997}.

Belief sets can be extracted from epistemic states (the belief set extracted from an epistemic state $\mathbb{E}$ is denoted as $B(\mathbb{E})$). Darwiche and Pearl show that their version of revision of epistemic states can be modelled with a tpo $\leq_{\mathbb{E}}$, associated with $\mathbb{E}$. Extraction of $B(\mathbb{E})$ is done by considering the lowest ranked worlds (i.e. the most plausible interpretations) under $\leq_{\mathbb{E}}$. The set of propositional sentences that holds in all those worlds is assumed to be the belief set of the agent. For notation let $min(\alpha, \leq_{\mathbb{E}})$ denote the set of minimal models for the propositional formula $\alpha$ under $\leq_{\mathbb{E}}$. Then $[B(\mathbb{E})] = min(\top, \leq_{\mathbb{E}})$ is the set of worlds that are models of the belief set $B(\mathbb{E})$ associated with $\mathbb{E}$.

\begin{example}
    Continuing example \ref{example:example-introduction} it is possible to model an epistemic state $\mathbb{E}$ with the tpo $\leq_{\mathbb{E}}$.

    $min(\top, \leq_{\mathbb{E}})$ is the set of worlds on the lowest rank, $\{010, 100\}$. A possible belief set is $B(\mathbb{E}) = \{p \vee q, \neg p \vee \neg q, \neg r\}$ as $[B(\mathbb{E})] = \{010, 100\} = min(\top, \leq_{\mathbb{E}})$. Intuitively this makes sense with the initial assumption that John or Mary are the suspects ($p \vee q$), have acted alone ($\neg p \vee \neg q$) and the victim is not an alien ($\neg r$).
\end{example}

% TODO: do we need to discuss this in ore detail? faithful assignment here:
%Katsuno and Mendelzon [17] propose that an epistemic state Ψ
%should be equipped with an ordering ≤Ψ of the worlds (interpretations), where the compatibility with Bel(Ψ) is ensured by the socalled faithfulness.
%Definition 1 (Faithful Assignment [17]) A function Ψ 
%→≤Ψ that
%maps each epistemic state to a total preorder on interpretations is
%said to be a faithful assignment if and only if:
%(FA1) if ω1 ∈ Ψ and ω2 ∈ Ψ, then ω1 Ψ ω2
%(FA2) if ω1 ∈ Ψ and ω2 ∈  / Ψ, then ω1 <Ψ ω2
%Intuitively, ≤Ψ orders the worlds by plausibility, such that the minimal worlds with respect to ≤Ψ are the most plausible worlds.

\subsection{DP-AGM postulates}
\label{chapter:dp-agm-postulates}
The postulates first introduced by \cite{Alchourron1985} and reformulated for iterated belief revision on epistemic states by \cite{Darwiche1997} are repeated here. $\mathbb{E}$ denotes an epistemic state, $B(\mathbb{E})$ it's associated belief set, $B(\mathbb{E}) + \alpha$ the expansion of $B(\mathbb{E})$ by $\alpha$ and $\ast$ being a belief change operator on epistemic states.

\begin{enumerate}[wide=0pt, widest=99,leftmargin=\parindent,label = ($\mathbb{E}\!*\!\arabic*$)]
    \item\label{E1} $\qquad B(\mathbb{E}\ast\alpha) = Cn(B(\mathbb{E}\ast\alpha))$
    \item\label{E2} $\qquad \alpha \in B(\mathbb{E}\ast\alpha)$
    \item\label{E3} $\qquad B(\mathbb{E}\ast\alpha)  \subseteq B(\mathbb{E})+\alpha$
    \item\label{E4} $\qquad \textrm{If } \alpha \notin B(\mathbb{E}) \textrm{ then } B(\mathbb{E}) + \alpha \subseteq B(\mathbb{E} \ast \alpha)$
    \item\label{E5} $\qquad \textrm{If } \mathbb{E} = \mathbb{F} \textrm{ and } \alpha \equiv \beta \textrm{ then } B(\mathbb{E} \ast \alpha) = B(\mathbb{F} \ast \beta)$
    \item\label{E6} $\qquad \bot \in B(\mathbb{E} \ast \alpha) \textrm{ iff } \vdash \neg \alpha$
    \item\label{E7} $\qquad B(\mathbb{E} \ast (\alpha \wedge \beta)) \subseteq B(\mathbb{E} \ast \alpha) + \beta$
    \item\label{E8} $\qquad \textrm{If } \neg \beta \notin B(\mathbb{E} \ast \alpha) \textrm{ then } B(\mathbb{E} \ast \alpha) + \beta \subseteq B(\mathbb{E} \ast (\alpha \wedge \beta))$
\end{enumerate}

To guarantee the ability to extract unique belief sets from epistemic states after a revision by $\alpha$, Booth and Meyer \cite{Booth2011} modify these postulates by requiring consistent epistemic inputs. That requirement means dropping \ref{E6} and considering only \ref{E1}-\ref{E5} and \ref{E7}-\ref{E8} as DP-AGM.

While DP-AGM define the most plausible worlds after a revision by $\alpha$ (as $min(\alpha, \leq_{\mathbb{E}})$), Darwiche and Pearl also define postulates that restrict the rest of the new ordering. Because the focus of this report is on the revision of tpos, the semantic (i.e. in terms of how the ordering of worlds undergoes change) versions are shown here, for more details and sentential versions refer to \cite{Darwiche1997}.

\begin{enumerate}[wide=0pt, widest=99,leftmargin=\parindent,label = (CR$\arabic*$)]
    \item\label{CR1} $\qquad \textrm{If } v\in [\alpha], w \in [\alpha] \textrm{ then } v \leq_{\mathbb{E}} w \textrm{ iff } v \leq_{\mathbb{E\ast\alpha}} w$
    \item\label{CR2} $\qquad \textrm{If } v\in [\neg \alpha], w \in [\neg \alpha] \textrm{ then } v \leq_{\mathbb{E}} w \textrm{ iff } v \leq_{\mathbb{E\ast\alpha}} w$
    \item\label{CR3} $\qquad \textrm{If } v\in [ \alpha], w \in [\neg \alpha] \textrm{ then } v <_{\mathbb{E}} w \textrm{ only if } v <_{\mathbb{E\ast\alpha}} w$
    \item\label{CR4} $\qquad \textrm{If } v\in [ \alpha], w \in [\neg \alpha] \textrm{ then } v \leq_{\mathbb{E}} w \textrm{ only if } v \leq_{\mathbb{E\ast\alpha}} w$
\end{enumerate}

\ref{CR1} and \ref{CR2} mean that the relative ordering of worlds following a revision by $\alpha$ stays the same, if the worlds are either both $\alpha$ or $\neg\alpha$ worlds. \ref{CR3} and \ref{CR4} require that $\alpha$-worlds that are strictly/weakly more plausible than $\neg\alpha$-worlds are still strictly/weakly more plausible than them after a $\alpha$-revision.

% TODO: QUESTION: do we need C1-C4 here? 
%\begin{equation*}
%\begin{aligned}
%    (C1) \qquad & \textrm{If } \beta \vdash \alpha \textrm{ then } B(\mathbb{E}\ast\alpha\ast\beta)=B(\mathbb{E}\ast\beta) \\
%    (C2) \qquad & \textrm{If } \beta \vdash \neg  \alpha \textrm{ then } B(\mathbb{E}\ast\alpha\ast\beta)=B(\mathbb{E}\ast\beta) \\
%    (C3) \qquad & \textrm{If } \alpha \in B(\mathbb{E}\ast\beta) \textrm{ then } \alpha \in B(\mathbb{E}\ast\alpha\ast\beta) \\
%    (C4) \qquad & \textrm{If } \neg \alpha \notin B(\mathbb{E}\ast\beta) \textrm{ then } \neg \alpha \notin B(\mathbb{E}\ast\alpha\ast\beta) \\
%\end{aligned}
%\end{equation*}

\section{Additional metadata for tpo revision}
\subsection{Enriching epistemic states}
Based on DP-AGM and the postulates $(CR1)$-$(CR4)$ Booth and Meyer assume a fixed tpo $\leq$ over a set of worlds $W$ that acts as a plausibility ordering. The goal of the paper is the discussion of functions $\ast$ that return a new ordering $\leq_{\alpha}^{\ast}$ for every $\alpha \in L$. These functions are referred to as revision operators for $\leq$.

% TODO this describes a concrete operator, not the idea. go back to more basics, w+ as positive representation, w- as negative. 
% TODO this also is maybe wrong? "add additional metadata to world w" is not what is happening, it is an additional structure W+-
Booth and Meyer's \cite{Booth2011} approach is to add additional metadata to every world $w \in W$. This new structure is denoted as $W^{\pm} = \{x^{\epsilon} \mid x \in W \textrm{ and } \epsilon \in \{+, -\}\}$. In this notation a world $w \in W$ is represented twice: When new evidence $\alpha$ arrives that makes $w$ more plausible (because $w \in [\alpha]$) the agent can assume it's rank as $w^{+} \in W^{\pm}$. In contrast if the new evidence makes $w$ less plausible ($w \notin [\alpha]$) then $w^{-} \in W^{\pm}$ is $w$'s rank. In their paper Booth and Meyer call the pair $(w^{+}, w^{-})$ the positive/negative representation of the world $w$. They are an abstract interval, representing metadata about plausibility assumptions about $w$.

\subsection{$\leq$-faithful tpos}
For the ordering on $W^{\pm}$ Booth and Meyer suppose an additional relation denoted $\preceq$ over $W^{\pm}$ that is added to the epistemic state of an agent.

\begin{example}
    \label{example:example-stick}
    Recall that tpos can be equivalently thought of as a linearly ordered set of ranks (Example \ref{example:example-introduction}, page \pageref{example:example-introduction}) and therefor a tpo over $W$ can be visualised by ordering every $w \in W$ in a table of ranks (Table \ref{tab:visualising-a-tpo-example}, page \pageref{tab:visualising-a-tpo-example}). To visualise the interval ordering introduced by $\preceq$, Booth and Meyer use a "stick" with the endpoints defined as $w^{+}$/$w^{-}$ respectively. This method was first introduced in \cite{Booth2007} and is shown in Figure \ref{fig:example-visualisation-scatterplot}.

    \begin{figure}[h]
        \centering
        \begin{tikzpicture}[scale=1.2]
            \begin{axis}[
                  yticklabels={,,$x_{4}$,$x_{3}$,$x_{2}$,$x_{1}$},
                  xticklabels={,,},
                  ytick style={draw=none},
                  xtick style={draw=none},
                  axis line style={draw=none}
            ]
            \addplot[
                scatter,
                scatter src=explicit symbolic,
                mark size=3,
                scatter/classes={
                    empty={mark=*, fill=white},
                    filled={mark=*, fill=black}
                },
                nodes near coords*={\Label},
                visualization depends on={value \thisrow{label} \as \Label}
            ]
            table [meta=class] {
                    x y class label

                    3 1 empty $x_{4}^{+}$
                    6 1 empty $x_{4}^{-}$

                    2 2 empty $x_{3}^{+}$
                    5 2 empty $x_{3}^{-}$

                    1 3 empty $x_{2}^{+}$
                    3 3 empty $x_{2}^{-}$

                    1 4 empty $x_{1}^{+}$
                    3 4 empty $x_{1}^{-}$
                    };
            \end{axis}
        \end{tikzpicture}
        \caption{Representation of $\preceq$ over $W^{\pm}$ using sticks}
        \label{fig:example-visualisation-scatterplot}
    \end{figure}

   The worlds on lower ranks (more on the left) are preferred and therefor assumed to be more plausible. Here for example $x_{1}^{+} \prec x_{3}^{+}$ and $x_{1}^{-} \sim x_{4}^{+}$ Note that while all sticks have the same length in this example that is not required.
\end{example}

To characterise the relations between $\preceq$ and $\leq$ Booth and Meyer define a list of conditions.

\begin{enumerate}[wide=0pt, widest=99,leftmargin=\parindent,label = ($\preceq\arabic*$)]
    \item\label{PREQ1} $\qquad \preceq \textrm{ is a tpo over } W^{\pm}$
    \item\label{PREQ2} $\qquad x^{+} \preceq y^{+} \textrm{ iff } x \leq y$
    \item\label{PREQ3} $\qquad x^{-} \preceq y^{-} \textrm{ iff } x \leq y$
    \item\label{PREQ4} $\qquad x^{+} \prec x^{-}$
\end{enumerate}

The choice between positive or negative representations of two worlds should be the same as under $\leq$ (due to \ref{PREQ2}, \ref{PREQ3}). According to \ref{PREQ4} there has to be a difference between two ranks of a positive representation and a negative representation of the same world. Given the choice between both, the positive representation has to be chosen.

\begin{definition}[$\leq$-faithful tpo over $W^{\pm}$ taken directly from \cite{Booth2011}]
\label{definition:faithful-tpo}Let $\preceq \in W^{\pm} \times W^{\pm}$. If $\preceq$ satisfies \ref{PREQ1}-\ref{PREQ4} we say $\preceq$ is a $\leq$-faithful tpo (over $W^{\pm}$).
\end{definition}

Because of \ref{PREQ2} and \ref{PREQ3} it is sufficient to only include $\preceq$ in the epistemic state. The tpo $\leq$ over $W$ can be determined from $\preceq$ by restricting to only $\{ x^{+} \mid x \in W\}$ or $\{ x^{-} \mid x \in W\}$ respectively.

\begin{example}
    \label{example:example-faithful-tpo}
    Is the following tpo $\preceq$ over $W^{\pm}$ a $\leq$-faithful tpo?

    \begin{figure}[h]
        \centering
        \begin{tikzpicture}[scale=1]
            \begin{axis}[
                  yticklabels={,,},
                  xticklabels={,,},
                  ytick style={draw=none},
                  xtick style={draw=none},
                  axis line style={draw=none}
            ]
            \addplot[
                scatter,
                scatter src=explicit symbolic,
                mark size=3,
                scatter/classes={
                    empty={mark=*, fill=white},
                    filled={mark=*, fill=black}
                },
                nodes near coords*={\Label},
                visualization depends on={value \thisrow{label} \as \Label}
            ]
            table [meta=class] {
                    x y class label

                    3 1 empty $x_{3}^{+}$
                    5 1 empty $x_{3}^{-}$

                    2 2 empty $x_{2}^{+}$
                    3 2 empty $x_{2}^{-}$

                    1 3 empty $x_{1}^{+}$
                    4 3 empty $x_{1}^{-}$
                    };
            \end{axis}
        \end{tikzpicture}
    \end{figure}

    While \ref{PREQ1} and \ref{PREQ4} are satisfied, $\preceq$ is not a $\leq$-faithful tpo: Due to \ref{PREQ2} and $x_{1}^{+} \prec x_{2}^{+}$ follows $x_{1} < x_{2}$. That means \ref{PREQ3} can not hold as $x_{2}^{-} \prec x_{1}^{-}$ which would require $x_{2} < x_{3}$ to be true.

    As demonstrated here while the "sticks" do not need to be the same size (they just must exist due to \ref{PREQ4}), they need to be the same size for worlds that share a rank for their corresponding representations. They can not "cover" other sticks completely to continue to satisfy \ref{PREQ2} and \ref{PREQ3}.
\end{example}

\section{Booth and Meyer tpo revision operators}
Booth and Meyer \cite{Booth2011} aim to discuss tpo revision operators, functions $\ast$ that return a new tpo $\leq_{\alpha}^{\ast}$ for every $\alpha \in L$. To do so they define how to use the additional information from a $\leq$-faithful tpo $\preceq$ over $W^{\pm}$ to create a revision operator $\ast_{\preceq}$.

\begin{definition}[Revision operator $\ast_{\preceq}$ for $\leq$ generated by $\preceq$ from \cite{Booth2011}]
    \label{definition:revision-operator}
    For each $\leq$-faithful tpo $\preceq$ over $W^{\pm}$, refer to $\ast_{\preceq}$ as the revision operator for $\leq$ generated by $\preceq$ defined by:

    Set for any $\alpha \in L$ and $x \in W$:
    \begin{equation*}
        r_{\alpha}(x) = \left\{
                    \begin{array}{ll}
                      x^{+} \textrm{ if } x \in [\alpha]\\
                      x^{-} \textrm{ if } x \in [\neg\alpha]
                    \end{array}
                  \right.
    \end{equation*}

    The revised tpo $\leq_{\alpha}^{\ast}$ is defined by setting, for each $x, y \in W$,

    \begin{equation*}
        x \leq_{\alpha}^{\ast} y \textrm{ iff } r_{\alpha}(x) \preceq r_{\alpha}(y)
    \end{equation*}
\end{definition}

Intuitively new evidence $\alpha$ makes worlds that satisfy $\alpha$ more plausible and worlds that do not satisfy $\alpha$ less plausible. Therefor the agent associates worlds $x \in [\alpha]$ with their positive representation $x^{+}$ and worlds $x \in [\neg\alpha]$ with their negative representation $x^{-}$.

\begin{example}
    \label{example:example-updated-tpo}
    Reconsider the $\leq$-faithful tpo $\preceq$ from example \ref{example:example-stick} (page \pageref{example:example-stick}). When a new piece of evidence $\alpha$ has to be considered each world gets mapped to one end of the interval assigned to it, depending on if it is an $\alpha$-world or not. In figure \ref{fig:example-updated-tpo} that mapping is indicated by the filled out dot. For this example $x_{1}, x_{2}$ and $x_{3}$ are $[\neg\alpha]$-worlds while $x_{4}$ is an $[\alpha]$-world.

    \begin{figure}[h]
        \centering
        \begin{tikzpicture}[scale=1.2]
            \begin{axis}[
                  yticklabels={,,$x_{4}$,$x_{3}$,$x_{2}$,$x_{1}$},
                  xticklabels={,,},
                  ytick style={draw=none},
                  xtick style={draw=none},
                  axis line style={draw=none}
            ]
            \addplot[
                scatter,
                scatter src=explicit symbolic,
                mark size=3,
                scatter/classes={
                    empty={mark=*, fill=white},
                    filled={mark=*, fill=black}
                },
                nodes near coords*={\Label},
                visualization depends on={value \thisrow{label} \as \Label}
            ]
            table [meta=class] {
                    x y class label

                    3 1 filled $x_{4}^{+}$
                    6 1 empty $x_{4}^{-}$

                    2 2 empty $x_{3}^{+}$
                    5 2 filled $x_{3}^{-}$

                    1 3 empty $x_{2}^{+}$
                    3 3 filled $x_{2}^{-}$

                    1 4 empty $x_{1}^{+}$
                    3 4 filled $x_{1}^{-}$
                    };
            \end{axis}
        \end{tikzpicture}
        \caption{Associating worlds with their positive/negative representations}
        \label{fig:example-updated-tpo}
    \end{figure}

    The updated tpo $\leq_{\alpha}^{\ast}$ can now be inferred from the figure. It is the ordering of the filled dots, representing the new assignment for the respective world: $x_{1} \sim x_{2} \sim x_{4} < x_{3}$.
\end{example}

\subsection{Non-prioritised revision}

Belief revision with the AGM postulates \cite{Alchourron1985} or the reformulation by Darwiche and Pearl \cite{Darwiche1997} always includes the new information $\alpha$ in the belief set after revision. This feature is explicit in \ref{E2} of DP-AGM from chapter \ref{chapter:dp-agm-postulates} (page \pageref{chapter:dp-agm-postulates}), given as $\alpha \in B(\mathbb{E}\ast\alpha)$.

The operators characterised by Booth and Meyer do not require new evidence $\alpha$ to be part of the revised belief set, which makes them part of non-prioritised revision operators \cite{Hansson1999}.

A demonstration is already provided with example \ref{example:example-updated-tpo} and figure \ref{fig:example-updated-tpo}. The belief set extracted from the new epistemic state $\mathbb{E}$ as $[B(\mathbb{E})] = min(\top, \leq_{\alpha}^{\ast})$ includes the $[\neg\alpha]$-worlds $x_{1}$ and $x_{2}$ and therefor does not include the new evidence $\alpha$.

%Although the approach described allows us to determine more than just
%the belief set associated with an epistemic state, there is a problem with this
%approach regarding iterated tpo-revision. While the extra structure tells us
%how to determine a new tpo, it tells us nothing about how to determine the
%new extra structure which is needed to guide the next revision. Clearly the
%problem of iterated belief revision has simply re-emerged “one level up”. We
%investigate this problem by considering some desirable properties for revising
%the extra structure, and prove the consistency of these properties by giving a
%concrete operator satisfying all of them.

% TODO write Properties of Booth and Meyer tpo-revision operators
\section{Properties of Booth and Meyer tpo-revision operators}
While not determining a concrete operator $\ast_{\preceq}$ to revise a tpo $\leq$, definition \ref{definition:revision-operator} allows Booth and Meyer to discuss what properties any operator in that family must have \cite{Booth2011}.

The list of properties they consider to be a complete axiomatisation of the family of operators they describe follows:

\begin{enumerate}[wide=0pt, widest=99,leftmargin=\parindent,label = ($\ast\arabic*$)]
    \item\label{AST1} $\qquad \leq_{\alpha}^{\ast} \textrm{ is a tpo over } W$
    \item\label{AST2} $\qquad\alpha \equiv \gamma \textrm{ implies } \leq_{\alpha}^{\ast}=\leq_{\gamma}^{\ast}$
    \item\label{AST3} $\qquad \textrm{If } x, y \in [\alpha] \textrm{ then } x \leq_{\alpha}^{\ast} y \textrm{ iff } x \leq y$
    \item\label{AST4} $\qquad \textrm{If } x, y \in [\neg\alpha] \textrm{ then } x \leq_{\alpha}^{\ast} y \textrm{ iff } x \leq y$
    \item\label{AST5} $\qquad \textrm{If } x \in [\alpha], y \in [\neg\alpha] \textrm{ and } x \leq y \textrm{ then } x <_{\alpha}^{\ast} y$
    \item\label{AST6} $\qquad \textrm{If } x \in [\alpha], y \in [\neg\alpha] \textrm{ and } y \leq_{\alpha}^{\ast} x \textrm{ then } y \leq_{\gamma}^{\ast} x$
    \item\label{AST7} $\qquad \textrm{If } x \in [\alpha], y \in [\neg\alpha] \textrm{ and } y <_{\alpha}^{\ast} x \textrm{ then } y <_{\gamma}^{\ast} x$
\end{enumerate}

%TODO question: the text sounds like the properties are supposed by BM/they define them but they are also proved?! or does this mean proved for their approach but could be done differently?
%We propose that if x is not more
%preferred than y, even after receiving evidence which clearly points more to x
%being the case than it does to y, then there can be no evi

\ref{AST1} and \ref{AST2} are considered basic properties: A revision of a tpo $\leq$ over $W$ must generate another tpo over $W$ . In addition the operator must return an equal revised tpo for semantically equivalent sentences, the so called syntax-irrelevance property.

The next group of properties, \ref{AST3}-\ref{AST5} are common rules in iterated belief change. Booth and Meyer themselves consider them characteristic for admissible revision operators \cite{Booth2006a}.
If revising a tpo by a sentence $\alpha$, the relative ordering of worlds, that are either both $[\alpha]$- or $[\neg\alpha]$-worlds, must stay the same (\ref{AST3}/\ref{AST4}). These properties were already part of the semantic postulates \ref{CR1} and \ref{CR2} by Darwiche and Pearl \cite{Darwiche1997} discussed in chapter \ref{chapter:dp-agm-postulates} (page \pageref{chapter:dp-agm-postulates}).
% TODO this formulation is a disaster
In a similar way \ref{AST5} is a stronger requirement than the other two Darwiche and Pearl postulates \ref{CR3} and \ref{CR4}. It means the following: A world $x$ that is considered at least as plausible as a world $y$, before a revision by new information $\alpha$. If $\alpha$ makes $x$ more plausible ($x \in [\alpha]$) and $y$ less plausible ($y \in [\neg\alpha]$) then after revision $x$ should be considered strictly more plausible than $y$. This rule was not only proposed by Booth and Meyer \cite{Booth2006a} but also Jin und Thielscher in \cite{Jin2007} (as the postulate of independence).

% TODO: WHAT ARE ADMISSIBLE OPERATORS
%info here: \cite{Booth2006a}

% TODO: pull these out? Basic postulates for tpo revision in one subsection, additional ones "supplementary rationality properties" as second section?
These properties define tpo-revision with one input sentence, which make Booth and Meyer relate them to the AGM postulates for belief set revison \cite{Alchourron1985} as basic postulates for tpo-revision \cite{Booth2011}.

With \ref{AST6} and \ref{AST7} Booth and Meyer aim to add supplementary rationality properties to keep revision with different sentences coherent. These properties are largely ignored in other literature on iterated belief change \cite{Booth2011}. \ref{AST6} means that if, after revision by evidence $\alpha$ that makes a world $x$ more plausible than $y$, $y$ is still considered at least as plausible as $x$, then for any possible revision by evidence $\gamma$, $y$ must be at least as plausible as $x$. \ref{AST7} is the equivalent property for strict preference.

\bigskip

With these properties for $\ast$ defined, Booth and Meyer define the family of operators they are discussing with theorem \ref{theorem:revision-operator}.

% THEOREM 1 as axiomatisation of the family of revision operators considered in the paper. "It turns out that these properties provide an exact characterisation of the
% revision operators we consider."

\begin{theorem}
\label{theorem:revision-operator}Let $\ast$ be any revision operator for $\leq$. Then $\ast$ is generated from some $\leq$-faithful tpo $\preceq$ over $W^{\pm}$ iff $\ast$ satisfies \ref{AST1}-\ref{AST7}. \cite{Booth2011}
\end{theorem}

% TODO: do example operator for tpo <= here for each property with alien example and proof it works for each? -> OWN contrib

% TODO write Sentinential view
\subsection{Sentinential view}
% booth: "on a semantic level, i.e., in terms of how the ordering of individual worlds x, y undergo change"
%    \item Explanation of sentinential view (and distinction to semantic level), definition of $\leq' \circ \beta $ and how it relates to conditional beliefs, mention of $\leq' \circ \top $ as belief set. Show of properties for the family of revision operators using sentinential view. Explanation of Disj1/Disj2 as properties that are easy to show in the sentinential view but not in the semantic formulation.
%\end{itemize}
%\textbf{Section Goal:} Definition of $\ast_{\preceq}$ as example revision operator and visualisation of it. Show of axiomatic definition of the family of operators in semantic and sentinential view. Establishing theorem 1.

\section{Revising $\preceq$}
% DO NOT DO THIS? OR mention that we only show a concrete operator and not discuss properties, see discussion
\subsection{Example operator}
% shortened example from paper page 232 without SPH teil, show that shortened example satisfies the properties later on. USE EXAMPLE from darwiche -> here to illustrate things with guy killing an alien
%TODO this could be OWN contribution!!?

% TODO write Discussion
\section{Discussion}
% TODO *6,*7 seem weird, no different "strength" of evidence?
% TODO: discuss 6/7 being wrong? strength of evidence?=!=!, discuss 5 with it being equal when sticks - and + are same rank?! 
\subsection{Iterating on the iterator}
The revision operator $\ast$ based on $\preceq$ allows the revision of a tpo $\leq$ associated with an epistemic state to $\leq_{\alpha}^{\ast}$. To be useful in iterated belief revision the newly introduced structure $\preceq$ also needs to be revised, as Booth and Meyer put it the problem has "re-emerged 'one level up'" \cite{Booth2011}. In their paper Booth and Meyer discuss strict preference hierarchies (or SPHs), an equivalent structure to $\preceq$ denoted as $\$$, their properties and how to revise them. For more information on SPH revision see \cite{Booth2011}.

\newpage

\typeout{}
\bibliographystyle{plain}
\bibliography{references}

\newpage

\section*{Erklärung}
Ich erkläre, dass ich die schriftliche Ausarbeitung zum Seminar selbstständig und ohne unzulässige Inanspruchnahme Dritter verfasst habe. Ich habe dabei nur die angegebenen Quellen und Hilfsmittel verwendet und die aus diesen wörtlich oder sinngemäß entnommenen Stellen als solche kenntlich gemacht. Die Versicherung selbstständiger Arbeit gilt auch für enthaltene Zeichnungen, Skizzen oder graphische Darstellungen. Die Ausarbeitung wurde bisher in gleicher oder Ähnlicher Form weder derselben noch einer anderen Prüfungsbehörde vorgelegt und auch nicht veröffentlicht. Mit der Abgabe der elektronischen Fassung der endgültigen Version der Ausarbeitung nehme ich zur Kenntnis, dass diese mit Hilfe eines Plagiatserkennungsdienstes auf enthaltene Plagiate geprüft werden kann und ausschließlich für Prüfungszwecke gespeichert wird.

\end{document}