### 2021-03-08 presentation/1901_3230880_philip_heltweg_presentation.tex
feat: more notes in presentation, agm postulates

\documentclass[12pt, notheorems]{beamer}
%\documentclass[12pt, notheorems, handout]{beamer}
\usepackage[utf8]{inputenc}
\usepackage{pgfpages}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{dsfont}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{float}
\usepackage{stmaryrd}

\usepackage{enumitem}

% TODO QUESTION: do we need to discuss this in ore detail? faithful assignment here:

% ANSWER: Das ist eine gute Frage. So wie ich das Paper von Booth und Meyer im
% Kopf habe werden dort die Ordnung mit epistemischen Zuständen
% gleichgesetzt (was Darwiche und Pearl ja nicht machen). Von daher halte
% ich das für ok es nicht im aller formalen Ausführlichkeit zu machen.
% Aber man könnte es im einen Satz erwähnen

% Representation theorem
% In mathematics, a representation theorem is a theorem that states that every abstract structure with certain properties is isomorphic to another (abstract or concrete) structure.
% In mathematics, an isomorphism is a structure-preserving mapping between two structures of the same type that can be reversed by an inverse mapping.

% Operationalization
% In research design, especially in psychology, social sciences, life sciences and physics, operationalization or operationalisation is a process of defining the measurement of a phenomenon that is not directly measurable, though its existence is inferred by other phenomena.
% In a broader sense, it defines the extension of a concept—describing what is and is not an instance of that concept.

\pgfplotsset{compat=1.17}

\setitemize{label=\usebeamerfont*{itemize item}
  \usebeamercolor[fg]{itemize item}
  \usebeamertemplate{itemize item}}

\newcommand{\modelsOf}[1]{\llbracket #1 \rrbracket}

\usefonttheme{professionalfonts}
\usetheme{Madrid}
\usecolortheme{orchid}

\setbeamertemplate{caption}[numbered]
\setbeamertemplate{theorems}[numbered]

% Unique numbering scheme for definitions and theorems with the "notheorems" options on loading beamer
% see https://tex.stackexchange.com/questions/82415/beamer-different-numbering-for-theorems-examples-definition-and-lemma
\makeatletter
    \ifbeamer@countsect
      \newtheorem{theorem}{\translate{Theorem}}[section]
    \else
      \newtheorem{theorem}{\translate{Theorem}}
    \fi
    \newtheorem{corollary}{\translate{Corollary}}
    \newtheorem{fact}{\translate{Fact}}
    \newtheorem{lemma}{\translate{Lemma}}
    \newtheorem{problem}{\translate{Problem}}
    \newtheorem{solution}{\translate{Solution}}

    \theoremstyle{definition}
    \newtheorem{definition}{\translate{Definition}}
    \newtheorem{definitions}{\translate{Definitions}}

    \theoremstyle{example}
    \newtheorem{example}{\translate{Example}}
    \newtheorem{examples}{\translate{Examples}}

    % Compatibility
    \newtheorem{Beispiel}{Beispiel}
    \newtheorem{Beispiele}{Beispiele}
    \theoremstyle{plain}
    \newtheorem{Loesung}{L\"osung}
    \newtheorem{Satz}{Satz}
    \newtheorem{Folgerung}{Folgerung}
    \newtheorem{Fakt}{Fakt}
    \newenvironment{Beweis}{\begin{proof}[Beweis.]}{\end{proof}}
    \newenvironment{Lemma}{\begin{lemma}}{\end{lemma}}
    \newenvironment{Proof}{\begin{proof}}{\end{proof}}
    \newenvironment{Theorem}{\begin{theorem}}{\end{theorem}}
    \newenvironment{Problem}{\begin{problem}}{\end{problem}}
    \newenvironment{Corollary}{\begin{corollary}}{\end{corollary}}
    \newenvironment{Example}{\begin{example}}{\end{example}}
    \newenvironment{Examples}{\begin{examples}}{\end{examples}}
    \newenvironment{Definition}{\begin{definition}}{\end{definition}}
\makeatother

% Notes
% Talk using pympress: brew install pympress
% Talk command: pympress -t 40 -n right ~/Documents/master/master-seminar/presentation/1901_3230880_philip_heltweg_presentation.pdf
\setbeameroption{show notes on second screen=right} % Both

% Coloring note-pages
\setbeamertemplate{note page}{\pagecolor{yellow!15}\insertnote\vfill\insertframenumber}
\setbeamertemplate{note page}{
	\pagecolor{yellow!15}
	\medskip
	SLIDE: \insertframenumber\\
	TITLE: \insertframetitle,

	\insertnote
}

\AtBeginSection[]
{
  \begin{frame}
    \frametitle{Table of Contents}
    \tableofcontents[currentsection]
  \end{frame}
}

\begin{document}
\title{Of judges, aliens and total preorders}
\author[Heltweg, Philip]{Heltweg, Philip\\pheltweg@gmail.com}
\institute{University of Hagen}
\date{March 13th, 2021}

\frame{\titlepage}
\note{
    \begin{itemize}
        \item Folien sind nummeriert unten rechts, bei Fragen am besten Nummer aufschreiben
    \end{itemize}
}

\begin{frame}{Table of Contents}
    \tableofcontents

    \note{
        \begin{itemize}
            \item Einführung mit Motivation, Forschungskontext und Beispiel
            \item Hintergrund: Festlegen Notation, was sind TPOs überhaupt, vorherige Arbeit von Alchouron et al, Darwiche und Pearl
            \item Metadata: Wie bauen Booth und Meyer auf der vorherigen Arbeit auf, Erklärung des Ansatzes zur TPO revision
            \item BM operators: Wie benutzen BM die vorher besprochenen Metadaten für belief revision, Definition von Eigenschaften ihrer Operatoren
            \item Iteration: Ausblick, wenn wir tpos mit der neuen Struktur anpassen können müssen wir auch die neue Struktur anpassen, ansonsten kann man nicht mehrere neue Informationen einpflegen. Hier haben BM sehr viel mehr geschrieben als hier vorgestellt aber ein konkreter Operator wird hier vorgestellt.
        \end{itemize}
    }
\end{frame}

\section{Introduction}
\note{
    \begin{itemize}
        \item Jetzt: Verstehen der Forschungsfrage, Einordnung des Papiers in einen Forschungskontext und Aufsetzen des begleitenden Beispiels ohne Formalisierung
    \end{itemize}
}

\begin{frame}{Motivation}
    How should a judge change their worldview when presented with new information?
    \note{
        \begin{itemize}
            \item Beispiel Frage zur Motivation, wird unser Beispiel sein. Wir haben zwei Verdächtige und diverse Beweise und Indizien, wie bilden wir uns ein Bild über die Welt?
            \item Unvollständige Information: Kein gesichertes Wissen sondern Glauben und Annahmen aus Zeugenaussagen/Indizien
            \item Inkonsistente Information: Zeugen können lügen, Indizien falsch sein, Informationen können mit dem aktuellen Glauben inkonsistent sein
            \item Keine reine richtig/falsch Frage. Soll neuer Information immer geglaubt werden? Wie sollen Konflikte gelöst werden? Wie kann man abbilden, dass der Richter mit Indizien für einen Sachverhalt immer mehr bereit ist an diesen Sachverhalt zu glauben?
        \end{itemize}
    }
\end{frame}

\begin{frame}{Research Context}
    \begin{itemize}
        \item Philosophy and artificial intelligence \cite{Ferme2011}
        \item Belief change \cite{Darwiche1997}
              \begin{itemize}
                  \item Nonmonotonic logic
                  \item Probabilistic reasoning
                  \item Belief revision
              \end{itemize}
        \item One-step vs. iterated belief revision
    \end{itemize}

    \note{
        \begin{itemize}
            \item Belief change, adapting one's world view under imperfect information
            \item Agent can be human or machine -> major areas philosophy and AI
            \item Different approaches, nonmonotonic logic known from beierle courses (truth maintenance systems, default logic like reiter or pool), probabilistic reasoning (also known from beierle course 1845),
            \item Belief revision: operator based. Define and discuss operators that transition agent state from one to the next for new evidence
            \item One step vs iterated depending on how many consecutive new information it handels
            \item Booth and Meyer: iterated belief revision
        \end{itemize}
    }
\end{frame}

\begin{frame}{Different types of belief}
    \begin{itemize}
        \item Belief set \cite{Alchourron1985}
        \item Conditional beliefs \cite{Darwiche1997}
        \item Change of conditional beliefs \cite{Booth2011}
    \end{itemize}

    \note{
        \begin{itemize}
            \item Belief sets as what an agent currently accepts as true (AGM theory: How is a belief set defined and how does the belief set change with new evidence?)
            \item Conditional beliefs as what an agent is prepared to accept as true depending on future evidence? (Darwiche and Pearl: In addition to how does the belief set change, what is the "strategy" for future belief set change? How do conditional beliefs change?)
            \item Booth and Meyer: Change of conditional beliefs: If conditional beliefs are the strategy to change belief sets with future evidence, how can we change conditional beliefs with new evidence?
        \end{itemize}
    }
\end{frame}

\begin{frame}{Research question}
    \begin{itemize}
        \item "conditional belief" revision operators
        \item axiomatisation of a family of operators by defining properties
        \item discuss properties and define concrete example
    \end{itemize}

    \note{
        \begin{itemize}
            \item Conditional beliefs/Conditional belief revision: Indizien für Sachverhalt = Sachverhalt nicht direkt glauben aber bei mehr Indizien immer mehr daran glauben/immer mehr bereit sein es zu akzeptieren.
            \item paper goal/focus of this talk revision operators for conditional belief structure (not mention tpo yet)
            \item postulates for operators to restrict them, define a family of operators (axiomatisation) by doing that
            \item properties are not "correct" but has to be done to reasonably talk about those operators and narrow down possibilities, compare to other properties/edge cases to branch out
        \end{itemize}
    }
\end{frame}

\section{Formal Background}
\note{
    \begin{itemize}
        \item Vorher: Verstehen der Forschungsfrage, Einordnung des Papiers in einen Forschungskontext und Aufsetzen des begleitenden Beispiels
        \item Jetzt: Formalisierung des Beispiels, Etablieren was eine total preorder ist.
        \item Jetzt: Einführung von Formalisierungen zu Glaubensmengen/Annahmen und wie preorders dort benutzt werden
        \item Jetzt: Postulate um Änderungen dieser Glaubensmengen/Annahmen mit neuen Informationen einzuschränken
    \end{itemize}
}

\begin{frame}{Courtroom example \footnote{inspired by \cite{Booth2011} and \cite{Darwiche1997}}}
    \begin{itemize}
        \item The agent is a judge in a murder trial, "John" and "Mary" are suspects, the victim might be an alien
        \item $\Sigma = \{ p, q, r\}$
              \begin{itemize}
                  \item p = "John is the murderer"
                  \item q = "Mary is the murderer"
                  \item r = "The victim is an alien"
              \end{itemize}
        \item  $Int({\Sigma}) = W = \{ 000, 001, 010, 011, 100, 101, 110, 111\}$
    \end{itemize}

    \note{
        \begin{itemize}
            \item Konsistentes Beispiel in der Präsentation, gut aufpassen
            \item Beispiel inspiriert aus der Literatur
            \item Aussagenlogik, Aussagenvariablen p, q, r
            \item "Mögliche Welt" als aussagenlogische Interpretationen/Belegungen, damit 2 hoch 3 = 8 mögliche Alternativen hier
            \item Kodierung der Welt als Tupel von Wahrheitswerten (0 = false, 1 = true)
            \item Beispiel für 1-2 Welten durchgehen: 010 = Mary ist Mörderin, Opfer ist kein Alien, 111 = John und Mary sind die Mörder, haben ein Alien getötet
        \end{itemize}
    }
\end{frame}

\begin{frame}{Total preorders}
    \begin{itemize}
        \item Common tool to handle preference orderings over propositional worlds \cite{Booth2011}
        \item binary relation $\leq$, total, reflexive, transitive
        \item  $<$ strict, $\sim$ symmetric closure
    \end{itemize}
    \note{
        \begin{itemize}
            \item total (for all $x, y \in W$ either $x \leq y$ or $y \leq x$)
            \item reflexive ($x \leq x$ for all $x \in W$)
            \item transitive (if $x \leq y$ and $y \leq z$ then $x \leq z$)
            \item  $<$ strict part of $\leq$
            \item $\sim$ symmetric closure of $\leq$ (i.e. $x \sim y$ iff $x \leq y$ and $y \leq x$)
        \end{itemize}
    }
\end{frame}

\begin{frame}{Preorder example}
    \begin{itemize}
        \item Judge beliefs
              \begin{itemize}
                  \item Murderer probably acted alone but possible that they conspired
                  \item Unlikely, but not impossible, for the victim to be an alien
              \end{itemize}
        \item $\leq$ over $W$: $010 \sim 100 < 000 \sim 110 < 011 \sim 101 < 001 \sim 111$
    \end{itemize}
    \note{
        \begin{itemize}
            \item Erklärung TPO, ein Mörderer und kein Alien gleich plausibel, dann kein/beide, dann jeweils mit Alien
        \end{itemize}
    }
\end{frame}

\begin{frame}{Preorder visualisation}

    \begin{itemize}
        \item $[[x]]_{\sim} = \{y \mid y \sim x\}$
        \item $[[x]] \leq [[y]]$ iff $x \leq y$
    \end{itemize}

    \begin{table}[H]
        \centering
        \begin{tabular}{llll}
            $R_{1}$                                          & $R_{2}$                                         & $R_{3}$                                         & $R_{4}$ \\ \hline
            \multicolumn{1}{|l|}{\begin{tabular}[c]{@{}l@{}}$010$\\ $100$\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}$000$\\ $110$\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}$011$\\ $101$\end{tabular}} &
            \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}$001$\\ $111$\end{tabular}}                                                                                                                \\ \hline
        \end{tabular}
        \caption{Visualizing a tpo as a linearly ordered set of ranks, as done in \cite{booth2006bad}}.
        \label{tab:visualising-a-tpo-example}
    \end{table}

    \note{
        \begin{itemize}
            \item TPOs können visualisiert werden als geordnete Liste von Rängen auf denen die Welten sind
            \item Die Ränge enthalten jeweils Welten mit gleicher Plausibilität
            \item Mathematisch ausgedrückt sind sie die Äquivalenzklassen zur Symmetrie (teilt die Welten in disjunkte Teilmengen ein)
        \end{itemize}
    }
\end{frame}

\begin{frame}{Belief Sets}
    \begin{itemize}
        \item Set of propositions the agent accepts as true at any point in time \cite{Ferme2011}
        \item Deductively closed
        \item Possible for example: $Cn(\{ p \vee q, \neg(p \wedge q), \neg r \})$
    \end{itemize}
    % TODO: Cn here good?

    \note{
        \begin{itemize}
            \item Belief sets als aktuell akzeptierte Sätze eines Agenten
            \item Gegensatz zu conditional beliefs: Sätze im Belief set werden als war angenommen.
            \item Betrachtungsgegenstand von AGM-Theorie \cite{Alchourron1985} die Postulate definiert haben um Belief set revision einzuschränken, rationale revision, wir sprechen das allerdings später an weil wir die neu formulierten Postulates von \cite{Darwiche1997} nutzen
        \end{itemize}
    }
\end{frame}

\begin{frame}{Epistemic States}
    \begin{itemize}
        \item abstract entity $\mathbb{E}$ that contain all information an agent need for their reasoning \cite{Darwiche1997}
        \item strategy for reasoning can be modeled as tpo $\leq_{\mathbb{E}}$ over worlds
        \item belief sets $B(\mathbb{E})$ can be extracted from epistemic states as all sentences true in the most plausible worlds
              \begin{itemize}
                  \item set of minimal worlds in which $\alpha$ is true: $min(\alpha, \leq_{\mathbb{E}})$, set of most plausible worlds:  $min(\top, \leq_{\mathbb{E}})$
                  \item Set of worlds that are models of the belief set of the agent $\modelsOf{B(\mathbb{E})} = min(\top, \leq_{\mathbb{E}})$
                  \item All sentences true in those worlds: $Th(min(\top, \leq_{\mathbb{E}}))$
              \end{itemize}
    \end{itemize}

    \note{
        \begin{itemize}
            \item Darwiche and Pearl argumentieren das Belief sets nicht ausreichend sind. Die Postulate von AGM definieren nur wie sich belief sets ändern (principle of least change) aber conditional beliefs können sich ändern wie sie wollen. Epistemic states enthalten die Strategie des Agenten wie er seinen Glauben ändert, damit auch die conditional beliefs.
            \item ("commonly used" does not mean has to be used, abstract) - BM nehmen es aber an in ihrem paper
        \end{itemize}
    }
\end{frame}

\begin{frame}{Belief Set Revision Postulates\footnote{AGM postulates \cite{Alchourron1985}, reformulated for epistemic states by Darwiche and Pearl \cite{Darwiche1997}}}
    \begin{enumerate}[wide=0pt, widest=99,leftmargin=\parindent,label = ($\mathbb{E}\!*\!\arabic*$)]
        \item\label{E1} $\qquad B(\mathbb{E}\ast\alpha) = Cn(B(\mathbb{E}\ast\alpha))$
        \item\label{E2} $\qquad \alpha \in B(\mathbb{E}\ast\alpha)$
        \item\label{E3} $\qquad B(\mathbb{E}\ast\alpha)  \subseteq B(\mathbb{E})+\alpha$
        \item\label{E4} $\qquad \textrm{If } \neg \alpha \notin B(\mathbb{E}) \textrm{ then } B(\mathbb{E}) + \alpha \subseteq B(\mathbb{E} \ast \alpha)$
        \item\label{E5} $\qquad \textrm{If } \mathbb{E} = \mathbb{F} \textrm{ and } \alpha \equiv \beta \textrm{ then } B(\mathbb{E} \ast \alpha) = B(\mathbb{F} \ast \beta)$
        \item\label{E6} $\qquad \bot \in B(\mathbb{E} \ast \alpha) \textrm{ iff } \models \neg \alpha$
        \item\label{E7} $\qquad B(\mathbb{E} \ast (\alpha \wedge \beta)) \subseteq B(\mathbb{E} \ast \alpha) + \beta$
        \item\label{E8} $\qquad \textrm{If } \neg \beta \notin B(\mathbb{E} \ast \alpha) \textrm{ then } B(\mathbb{E} \ast \alpha) + \beta \subseteq B(\mathbb{E} \ast (\alpha \wedge \beta))$
    \end{enumerate}

    \note{
        \begin{itemize}
            \item $\mathbb{E}$ als epistemischer status, $B(\mathbb{E})$ belief set des status, $B(\mathbb{E}) + \alpha$ ist die expansion $B(\mathbb{E})$ mit $\alpha$ ($\alpha$ hinzufügen. $CN(B(\mathbb{E}) \cup \alpha)$), $\ast$ being a belief change operator on epistemic states
            \item Postulates
                  \begin{itemize}
                      \item 1: "Closure": deduktiv abgeschlossen, ist ein belief set
                      \item 2: "Success": neue information ist immer teil des belief sets nach revision. Das muss nicht so sein, für BM tpo revision operatoren die wir hier betrachten gilt dies z.b. nicht. Wird "non-priotized revision" genannt
                      \item 3 und 4: "Inclusion" und "Vacuity": zusammen, wenn neues belief ist konsistent mit dem vorherigen belief set dann gibt es keinen grund beliefs zu entfernen. das neue belief set wird vorheriges belief set + alpha + was aus alpha und belief set folgt sein, minimal change wenn alpha konsistent mit belief set
                      \item 5: Syntax irrelevanz für information mit der revision betrieben wird. Epistemische zustände müssen gleich sein hier, im AGM original muss nur das belief set der zustände gleich sein. Also ist E5 schwächer als das Original von AGM.
                      \item 6: Das belief set enthält elementaren widerspruch NUR nach revision mit inkonsistenten informationen, konsistenz "at all costs" (alpha wärend not alpha eine tautologie ist). BM bestehen auf konsistente Eingaben damit sie immer ein belief set aus dem epistemischen status ableiten können. Daher nehmen sie E6 nicht an und nur E1-5, E7-8 als "DP-AGM"
                      \item 7 und 8: Wenn ein belief set nach revision mit alpha konsistent ist mit beta dann kann kann man das belief set von alpha und beta durch expansion mit beta erreichen (revision nicht nötig), wieder minimal change
                  \end{itemize}
        \end{itemize}
    }
\end{frame}

\begin{frame}{Conditional Belief Revision Postulates\footnote{by Darwiche and Pearl \cite{Darwiche1997}}}
    \begin{enumerate}[wide=0pt, widest=99,leftmargin=\parindent,label = (CR$\arabic*$)]
        \item\label{CR1} $\qquad \textrm{If } v\in \modelsOf{\alpha}, w \in \modelsOf{\alpha} \textrm{ then } v \leq_{\mathbb{E}} w \textrm{ iff } v \leq_{\mathbb{E\ast\alpha}} w$
        \item\label{CR2} $\qquad \textrm{If } v\in \modelsOf{\neg\alpha}, w \in \modelsOf{\neg\alpha} \textrm{ then } v \leq_{\mathbb{E}} w \textrm{ iff } v \leq_{\mathbb{E\ast\alpha}} w$
        \item\label{CR3} $\qquad \textrm{If } v\in \modelsOf{\alpha}, w \in \modelsOf{\neg\alpha} \textrm{ then } v <_{\mathbb{E}} w \textrm{ only if } v <_{\mathbb{E\ast\alpha}} w$
        \item\label{CR4} $\qquad \textrm{If } v\in \modelsOf{\alpha}, w \in \modelsOf{\neg\alpha} \textrm{ then } v \leq_{\mathbb{E}} w \textrm{ only if } v \leq_{\mathbb{E\ast\alpha}} w$
    \end{enumerate}
    \note{
        \begin{itemize}
            \item Zusätzliche postulate von DP um conditional belief revision einzuschränken
            \item Semantische version aka wie sich die Ordnung der Welten ändert mit v,w in W
            \item CR1/2: Relative Ordnung der alpha/not alpha welten bleibt gleich bei alpha/not alpha revision
            \item CR3 wenn eine alpha welt strikt bevorzugt wurde vor einer not alpha welt dann muss das auch nach einer revision mit alpha gelten
            \item CR4 wie CR3 nur für "weak preference"
        \end{itemize}
    }
\end{frame}

\section{Additional metadata for tpo revision}
\note{
    \begin{itemize}
        \item Vorher: Formalisierung des Beispiels/Glaubensmengen/Epistemischen Zuständen, Etablieren was eine total preorder ist, Postulate für Revision von Glaubensmengen und Annahmen
        \item Jetzt: BM wollen revision von tpos besprechen und führen dafür neue Metadaten und Strukturen im epistemischen Zustand ein, wir stellen diese nun vor bevor wir im nächsten Schritt die Änderung selbst besprechen
        \item Jetzt: Einführung einer intuitiven Visualisierung der neuen Struktur die wir im weiteren Verlauf nutzen
    \end{itemize}
}

\begin{frame}{Enriching Epistemic States}
    \begin{itemize}
        \item Additional structure $W^{\pm} = \{x^{\epsilon} \mid x \in W \textrm{ and } \epsilon \in \{+, -\}\}$
        \item "A bad day surfing is better than a good day working" \cite{booth2006bad}
        \item Worlds are either supported by evidence or not $w \in \modelsOf{\alpha}$ / $w \in \modelsOf{\neg\alpha}$
        \item Interval representing a world $(w^{+}, w^{-})$
    \end{itemize}
    \note{
        \begin{itemize}
            \item Additional structure of plausibility of worlds under new evidence  W+-
            \item "A bad day surfing is better than a good day working" als idee das eine Welt sogar unter evidenz die sie nicht supported mehr plausibel bleiben kann, e.g. aliens tendenziell nicht so einfach zu glauben
        \end{itemize}
    }
\end{frame}

\begin{frame}{$\leq$-faithful tpo}
    \begin{itemize}
        \item original tpo $\leq$ was an order over $W$
        \item $\preceq$ over new $W^{\pm}$
    \end{itemize}

    \note{
        \begin{itemize}
            \item what is it
            \item say that the original tpo can be reconstructed from it
        \end{itemize}
    }
\end{frame}

\begin{frame}{$\leq$-faithful tpo - definition}
    \begin{enumerate}[wide=0pt, widest=99,leftmargin=\parindent,label = ($\preceq\arabic*$)]
        \item\label{PREQ1} $\qquad \preceq \textrm{ is a tpo over } W^{\pm}$
        \item\label{PREQ2} $\qquad x^{+} \preceq y^{+} \textrm{ iff } x \leq y$
        \item\label{PREQ3} $\qquad x^{-} \preceq y^{-} \textrm{ iff } x \leq y$
        \item\label{PREQ4} $\qquad x^{+} \prec x^{-}$
    \end{enumerate}

    \begin{definition}[$\leq$-faithful tpo over $W^{\pm}$ \cite{Booth2011}]
        \label{definition:faithful-tpo}Let $\preceq \subseteq W^{\pm} \times W^{\pm}$. If $\preceq$ satisfies \ref{PREQ1}-\ref{PREQ4}, we say $\preceq$ is a $\leq$-faithful tpo (over $W^{\pm}$).
    \end{definition}
    \note{
        \begin{itemize}
            \item characteristics + definition
            \item original tpo can be reconstructed from this
        \end{itemize}
    }
\end{frame}

\begin{frame}[fragile]{$\leq$-faithful tpo visualisation}
    \begin{figure}[H]
        \centering
        \begin{tikzpicture}[scale=1]
            \begin{axis}[
                    yticklabels={,,$x_{4}$,$x_{3}$,$x_{2}$,$x_{1}$},
                    xticklabels={,,},
                    ytick style={draw=none},
                    xtick style={draw=none},
                    axis line style={draw=none}
                ]
                \addplot[
                    scatter,
                    scatter src=explicit symbolic,
                    mark size=3,
                    scatter/classes={
                            empty={mark=*, fill=white},
                            filled={mark=*, fill=black}
                        },
                    nodes near coords*={\Label},
                    visualization depends on={value \thisrow{label} \as \Label}
                ]
                table [meta=class] {
                        x y class label

                        3 1 empty $x_{4}^{+}$
                        6 1 empty $x_{4}^{-}$

                        2 2 empty $x_{3}^{+}$
                        5 2 empty $x_{3}^{-}$

                        1 3 empty $x_{2}^{+}$
                        3 3 empty $x_{2}^{-}$

                        1 4 empty $x_{1}^{+}$
                        3 4 empty $x_{1}^{-}$
                    };
            \end{axis}
        \end{tikzpicture}
        \caption{Representation of $\preceq$ over $W^{\pm}$ using intervals}
        \label{fig:example-visualisation-scatterplot}
    \end{figure}

    \note{
        \begin{itemize}
            \item example with sticks
        \end{itemize}
    }
\end{frame}

\begin{frame}[fragile]{Courtroom example: $\leq$-faithful tpo}
    \begin{figure}[H]
        \centering
        \begin{tikzpicture}[scale=1]
            \begin{axis}[
                    xticklabels={},
                    yticklabels={},
                    extra y ticks={1, 2, 3, 4, 5, 6, 7, 8},
                    extra y tick labels={$111$, $001$, $101$, $011$, $110$, $000$, $100$, $010$},
                    ytick style={draw=none},
                    xtick style={draw=none},
                    axis line style={draw=none}
                ]
                \addplot[
                    scatter,
                    scatter src=explicit symbolic,
                    mark size=3,
                    scatter/classes={
                            empty={mark=*, fill=white},
                            filled={mark=*, fill=black}
                        },
                    nodes near coords*={\Label},
                    visualization depends on={value \thisrow{label} \as \Label}
                ]
                table [meta=class] {
                        x y class label

                        1 8 empty \empty
                        3 8 empty

                        1 7 empty
                        3 7 empty

                        2 6 empty
                        4 6 empty

                        2 5 empty
                        4 5 empty

                        5 4 empty
                        7 4 empty

                        5 3 empty
                        7 3 empty

                        6 2 empty
                        8 2 empty

                        6 1 empty
                        8 1 empty
                    };
            \end{axis}
        \end{tikzpicture}
        \caption{Representation of $\preceq$ over $W^{\pm}$ for the courtroom example}
        \label{fig:example-tpo-initial}
    \end{figure}

    \note{
        \begin{itemize}
            \item example for courtroom
        \end{itemize}
    }
\end{frame}

\section{Booth and Meyer tpo-revision operators}
\note{
    \begin{itemize}
        \item Vorher: Neue Metadaten/Strukturen im epistemischen Zustand und Visualisierung
        \item Jetzt: Nutzen dieser neuen Strukturen für tpo revision, Postulates für BM revision operatoren und Definition der Familie von BM revisons operatoren
        \item Jetzt: Abstrakte tpo revision am Beispiel
    \end{itemize}
}

\begin{frame}{BM tpo-revision operator}
    \begin{definition}[Revision operator $\ast_{\preceq}$ for $\leq$ generated by $\preceq$ \cite{Booth2011}]
        \label{definition:revision-operator}
        For each $\leq$-faithful tpo $\preceq$ over $W^{\pm}$, refer to $\ast_{\preceq}$ as the revision operator for $\leq$ generated by $\preceq$ defined by:

        Set for any $\alpha \in L$ and $x \in W$:
        \begin{equation*}
            r_{\alpha}(x) = \left\{
            \begin{array}{ll}
                x^{+} \textrm{ if } x \in \modelsOf{\alpha} \\
                x^{-} \textrm{ if } x \in \modelsOf{\neg\alpha}
            \end{array}
            \right.
        \end{equation*}

        The revised tpo $\leq_{\alpha}^{\ast}$ is defined by setting, for each $x, y \in W$,

        \begin{equation*}
            x \leq_{\alpha}^{\ast} y \textrm{ iff } r_{\alpha}(x) \preceq r_{\alpha}(y)
        \end{equation*}
    \end{definition}

    \note{
        \begin{itemize}
            \item definition
        \end{itemize}
    }
\end{frame}

\begin{frame}[fragile]{Courtroom example: Revision Visualised}

    \begin{figure}[H]
        \centering
        \begin{tikzpicture}[scale=1]
            \begin{axis}[
                    xticklabels={},
                    yticklabels={},
                    extra y ticks={1, 2, 3, 4, 5, 6, 7, 8},
                    extra y tick labels={$111$, $001$, $101$, $011$, $110$, $000$, $100$, $010$},
                    ytick style={draw=none},
                    xtick style={draw=none},
                    axis line style={draw=none}
                ]
                \addplot[
                    scatter,
                    scatter src=explicit symbolic,
                    mark size=3,
                    scatter/classes={
                            empty={mark=*, fill=white},
                            filled={mark=*, fill=black}
                        },
                    nodes near coords*={\Label},
                    visualization depends on={value \thisrow{label} \as \Label}
                ]
                table [meta=class] {
                        x y class label

                        1 8 empty \empty
                        3 8 filled

                        1 7 filled
                        3 7 empty

                        2 6 empty
                        4 6 filled

                        2 5 filled
                        4 5 empty

                        5 4 empty
                        7 4 filled

                        5 3 filled
                        7 3 empty

                        6 2 empty
                        8 2 filled

                        6 1 filled
                        8 1 empty
                    };
            \end{axis}
        \end{tikzpicture}
        \caption{Associating positive and negative representations of worlds after receiving evidence $\alpha=p$}
        \label{fig:example-tpo-revised}
    \end{figure}

    \note{
        \begin{itemize}
            \item example 7, figure 4
        \end{itemize}
    }
\end{frame}

\begin{frame}{Courtroom example: Revision 1}

    \begin{itemize}
        \item for $010, 110 \in W$, $010 \leq 110$ before, revise by $\alpha = p$
    \end{itemize}

    \begin{equation*}
        010 \in \modelsOf{\neg\alpha} \textrm{ : } r_{\alpha}(010) = 010^{-}
    \end{equation*}
    \begin{equation*}
        110 \in \modelsOf{\alpha} \textrm{ : } r_{\alpha}(110) = 110^{+}
    \end{equation*}

    \begin{itemize}
        \item $110^{+} \prec 010^{-}$ is true, set $110 <_{\alpha}^{\ast} 010$
        \item new tpo $\leq_{\alpha}^{\ast}$ is: $100 <_{\alpha}^{\ast} 110 <_{\alpha}^{\ast} 010 <_{\alpha}^{\ast} 000 <_{\alpha}^{\ast} 101 <_{\alpha}^{\ast} 111 <_{\alpha}^{\ast} 011 <_{\alpha}^{\ast} 001$
    \end{itemize}

    \note{
        \begin{itemize}
            \item example 7, math
        \end{itemize}
    }
\end{frame}

\begin{frame}{Courtroom example: Revision 2}
    \begin{itemize}
        \item new tpo $\leq_{\alpha}^{\ast}$ is: $100 <_{\alpha}^{\ast} 110 <_{\alpha}^{\ast} 010 <_{\alpha}^{\ast} 000 <_{\alpha}^{\ast} 101 <_{\alpha}^{\ast} 111 <_{\alpha}^{\ast} 011 <_{\alpha}^{\ast} 001$
        \item $min(\top, \leq_{\mathbb{E}}) = \{100\}$: "John is the murderer and the victim is not an alien".
        \item $\leq_{\alpha}^{\ast}$ as representation of the conditional beliefs
              \begin{itemize}
                  \item before $010 \leq 110$: "Both suspects being the murderer is less plausible than Mary being the murderer"
                  \item now $110 <_{\alpha}^{\ast} 010$: "Only Mary being the murderer less plausible than both conspiring".
              \end{itemize}
    \end{itemize}

    \note{
        \begin{itemize}
            \item example 7, math
        \end{itemize}
    }
\end{frame}

\begin{frame}{Properties of BM Revision Operators: Basic properties}
    \begin{enumerate}[wide=0pt, widest=99,leftmargin=\parindent,label = ($\ast\arabic*$)]
        \item\label{AST1} $\qquad \leq_{\alpha}^{\ast} \textrm{ is a tpo over } W$
        \item\label{AST2} $\qquad\alpha \equiv \gamma \textrm{ implies } \leq_{\alpha}^{\ast}=\leq_{\gamma}^{\ast}$
    \end{enumerate}

    \note{
        \begin{itemize}
            \item \ref{AST1} and \ref{AST2} are considered basic properties: A revision of a tpo $\leq$ over $W$ must generate another tpo over $W$. In addition, the operator must return an equal revised tpo for semantically equivalent sentences, the so called syntax-irrelevance property.
        \end{itemize}
    }
\end{frame}

\begin{frame}{Properties of BM Revision Operators: Common rules in iterated belief change}
    \begin{enumerate}[wide=0pt, widest=99,leftmargin=\parindent,label = ($\ast\arabic*$)]
        \setcounter{enumi}{2}
        \item\label{AST3} $\qquad \textrm{If } x, y \in \modelsOf{\alpha} \textrm{ then } x \leq_{\alpha}^{\ast} y \textrm{ iff } x \leq y$
        \item\label{AST4} $\qquad \textrm{If } x, y \in \modelsOf{\neg\alpha} \textrm{ then } x \leq_{\alpha}^{\ast} y \textrm{ iff } x \leq y$
        \item\label{AST5} $\qquad \textrm{If } x \in \modelsOf{\alpha}, y \in \modelsOf{\neg\alpha} \textrm{ and } x \leq y \textrm{ then } x <_{\alpha}^{\ast} y$
    \end{enumerate}

    \note{
        \begin{itemize}
            \item The next group of properties, \ref{AST3}-\ref{AST5} are common rules in iterated belief change. Booth and Meyer themselves consider them characteristic for admissible revision operators, a class of operators that requires new evidence to not be ignored completely \cite{Booth2006a}. This is in contrast to revision strategies like natural revision by Boutilier \cite{Boutilier1996}, which minimises change in conditional beliefs.
            \item If revising a tpo by a sentence $\alpha$, the relative ordering of worlds that are either both $\modelsOf{\alpha}$- or $\modelsOf{\neg\alpha}$-worlds, must stay the same (\ref{AST3}/\ref{AST4}). These properties were already part of the semantic postulates \ref{CR1} and \ref{CR2}, by Darwiche and Pearl \cite{Darwiche1997}, discussed in section \ref{chapter:dp-agm-postulates}.
            \item In a similar way, \ref{AST5} is a stronger requirement than the other two Darwiche and Pearl postulates \ref{CR3} and \ref{CR4}. It means the following: When a world $x$ is considered at least as plausible as a world $y$ before a revision by new information $\alpha$, if $\alpha$ makes $x$ more plausible ($x \in \modelsOf{\alpha}$) and $y$ less plausible ($y \in \modelsOf{\neg\alpha}$), then after revision by $\alpha$, $x$ should be considered strictly more plausible than $y$. This property was not only proposed by Booth and Meyer \cite{Booth2006a}, but also by Jin und Thielscher in \cite{Jin2007} (as the postulate of independence).
            \item These properties define tpo-revision with one input sentence. Booth and Meyer relate them to the AGM postulates for belief set revision \cite{Alchourron1985} as basic postulates for tpo-revision \cite{Booth2011}.
        \end{itemize}
    }
\end{frame}

\begin{frame}{Properties of BM Revision Operators: Supplementary rationality properties }
    \begin{enumerate}[wide=0pt, widest=99,leftmargin=\parindent,label = ($\ast\arabic*$)]
        \setcounter{enumi}{5}
        \item\label{AST6} $\qquad \textrm{If } x \in \modelsOf{\alpha}, y \in \modelsOf{\neg\alpha} \textrm{ and } y \leq_{\alpha}^{\ast} x \textrm{ then } y \leq_{\gamma}^{\ast} x$
        \item\label{AST7} $\qquad \textrm{If } x \in \modelsOf{\alpha}, y \in \modelsOf{\neg\alpha} \textrm{ and } y <_{\alpha}^{\ast} x \textrm{ then } y <_{\gamma}^{\ast} x$
    \end{enumerate}

    \note{
        \begin{itemize}
            \item With \ref{AST6} and \ref{AST7}, Booth and Meyer aim to add supplementary rationality properties to keep revision with different sentences coherent. These properties are largely ignored in other literature on iterated belief change \cite{Booth2011}. \ref{AST6} means that if, after revision by evidence $\alpha$ that makes a world $x$ more plausible than $y$, $y$ is still considered at least as plausible as $x$, then for revision by any possible evidence $\gamma$, $y$ must still be at least as plausible as $x$. \ref{AST7} is the equivalent property for strict preference.
                  %\item so far each of our rules mention onlyonerevision input sentenceα(modulo logical equivalence).  Byanalogy  with the AGM postulates  forbelief setrevision ,  we might consider them as the set ofbasicpostulates  for  tpo-revision.   One  thing  largely  missing  from  the  literature  on  iterated  belief  change  is  aserious study ofsupplementaryrationality properties which bestow a certain amount of coherence on theresults of revising bydifferentsentences.  The last couple of properties do this.
        \end{itemize}
    }
\end{frame}

\begin{frame}{Family of BM Revision Operators}
    \begin{theorem}
        \label{theorem:revision-operator}Let $\ast$ be any revision operator for $\leq$. Then $\ast$ is generated from some $\leq$-faithful tpo $\preceq$ over $W^{\pm}$ iff $\ast$ satisfies \ref{AST1}-\ref{AST7}. \cite{Booth2011}
    \end{theorem}

    \note{
        \begin{itemize}
            \item theorem 1
        \end{itemize}
    }
\end{frame}

\begin{frame}[fragile]{Non-priotized revision}
    \begin{figure}[H]
        \centering
        \begin{tikzpicture}[scale=1]
            \begin{axis}[
                    xticklabels={},
                    yticklabels={},
                    extra y ticks={1, 2, 3, 4, 5, 6, 7, 8},
                    extra y tick labels={$111$, $001$, $101$, $011$, $110$, $000$, $100$, $010$},
                    ytick style={draw=none},
                    xtick style={draw=none},
                    axis line style={draw=none}
                ]
                \addplot[
                    scatter,
                    scatter src=explicit symbolic,
                    mark size=3,
                    scatter/classes={
                            empty={mark=*, fill=white},
                            filled={mark=*, fill=black}
                        },
                    nodes near coords*={\Label},
                    visualization depends on={value \thisrow{label} \as \Label}
                ]
                table [meta=class] {
                        x y class label

                        1 8 empty \empty
                        3 8 filled

                        1 7 empty
                        3 7 filled

                        2 6 empty
                        4 6 filled

                        2 5 empty
                        4 5 filled

                        5 4 filled
                        7 4 empty

                        5 3 filled
                        7 3 empty

                        6 2 filled
                        8 2 empty

                        6 1 filled
                        8 1 empty
                    };
            \end{axis}
        \end{tikzpicture}
        \caption{Non-prioritised revision by $\alpha = r$}
        \label{fig:example-non-prio-revision}
    \end{figure}

    \note{
        \begin{itemize}
            \item Es gibt postulate die erfordern, dass neuen Informationen immer geglaubt wird und sie in das belief set aufgenommen werden, e.g. \ref{E2} als $\alpha \in B(\mathbb{E}\ast\alpha)$
            \item BM revision operatoren sind "non priotized revision", d.h. neue informationen müssen nicht teil des belief sets werden wie hier gezeigt
            \item aber auch erinnerung: admissible revision von BM erfordert (\ref{AST3} - \ref{AST5}), dass neue informationen nicht ignoriert werden, hier r verändert die conditional beliefs
            \item das im gegensatz zu natural revision by boutilier welches minimal change für conditional beliefs fordert
        \end{itemize}
    }
\end{frame}

\section{Iterating: \texorpdfstring{$\preceq$}{tpo}-revision}
\note{
    \begin{itemize}
        \item Vorher: BM revisions operatoren definition, Revision der tpo am Beispiel
        \item Jetzt: Revision der neuen Struktur $\preceq$, ansonsten Problem mehrmaliger Revision nur "hochgeschoben"
        \item Jetzt: Ein möglicher konkreter Beispieloperator + Visualisierung
        \item Jetzt: Wie die Revision von Annahmen funktioniert für mehr als einen Input
    \end{itemize}
}

\begin{frame}{A concrete operator \footnote{as shown in \cite{Booth2011}}: Setup}
    \begin{itemize}
        \item Function p mapping worlds to real numbers: $p: W^{\pm} \mapsto \mathds{R}$
        \item Interval representing a world: $(p(x^{+}), p(x^{-}))$
        \item Distance between representations: $p(x^{-}) - p(x^{+}) = a > 0$
        \item Define a tpo from $p$: $x^{\epsilon} \preceq_{p} y^{\delta} \textrm{ iff } p(x^{\epsilon}) \leq p(y^{\delta})$
    \end{itemize}

    \note{
        \begin{itemize}
            \item math for operator
        \end{itemize}
    }
\end{frame}

\begin{frame}{A concrete operator: Iteration}
    \begin{itemize}
        \item Choose initial $p$ so that $\preceq_{p} = \preceq$
        \item Revise $p$ by $\alpha$ to $p \ast \alpha$, for every $x^{\epsilon} \in W^{\pm}$:
    \end{itemize}

    \begin{equation*}
        (p \ast \alpha)(x^{\epsilon}) = \left\{
        \begin{array}{ll}
            p(x^{\epsilon}) \textrm{ if } x \in \modelsOf{\alpha} \\
            p(x^{\epsilon}) + a \textrm{ if } x \in \modelsOf{\neg\alpha}
        \end{array}
        \right.
    \end{equation*}

    \begin{itemize}
        \item Define a revised tpo $\preceq_{p \ast \alpha}$ from $p \ast \alpha$: $x^{\epsilon} \preceq_{p \ast \alpha} y^{\delta} \textrm{ iff } (p \ast \alpha)(x^{\epsilon}) \leq (p \ast \alpha)(y^{\delta})$
    \end{itemize}

    \note{
        \begin{itemize}
            \item math for operator
        \end{itemize}
    }
\end{frame}

\begin{frame}{Courtroom Example: A concrete operator}
    \begin{itemize}
        \item Choose initial $p$ so that $\preceq_{p} = \preceq$
              \begin{itemize}
                  \item $010 \textrm{ : } (p(010^{+}), p(010^{-})) = (0, a)$.
                  \item $100 \textrm{ : } (p(100^{+}), p(100^{-})) = (0, a)$.
              \end{itemize}
        \item Revise $p$ by $\alpha$ to $p \ast \alpha$
              \begin{itemize}
                  \item $010 \in \modelsOf{\neg\alpha} \textrm{ : } (p(010^{-}), p(010^{-}) + a) = (a, 2a)$
                  \item $100 \in \modelsOf{\alpha} \textrm{ : } (p(100^{+}), p(100^{-})) = (0, a)$
              \end{itemize}
    \end{itemize}

    \note{
        \begin{itemize}
            \item math for example with operator
        \end{itemize}
    }
\end{frame}

\begin{frame}[fragile]{Courtroom Example: Visualised}
    \begin{figure}[H]
        \centering
        \begin{tikzpicture}[scale=1]
            \begin{axis}[
                    xticklabels={},
                    yticklabels={},
                    extra y ticks={1, 2, 3, 4, 5, 6, 7, 8},
                    extra y tick labels={$111$, $001$, $101$, $011$, $110$, $000$, $100$, $010$},
                    ytick style={draw=none},
                    xtick style={draw=none},
                    axis line style={draw=none}
                ]
                \addplot[
                    scatter,
                    scatter src=explicit symbolic,
                    mark size=3,
                    scatter/classes={
                            empty={mark=*, fill=white},
                            old={mark=x, fill=white}
                        },
                    nodes near coords*={\Label},
                    visualization depends on={value \thisrow{label} \as \Label}
                ]
                table [meta=class] {
                        x y class label

                        1 8 old \empty

                        3 8 empty
                        5 8 empty

                        1 7 empty
                        3 7 empty

                        2 6 old

                        4 6 empty
                        6 6 empty

                        2 5 empty
                        4 5 empty

                        5 4 old

                        7 4 empty
                        9 4 empty

                        5 3 empty
                        7 3 empty

                        6 2 old

                        8 2 empty
                        10 2 empty

                        6 1 empty
                        8 1 empty
                    };
            \end{axis}
        \end{tikzpicture}
        \caption{$\preceq_{p \ast \alpha}$ for $\alpha = p$}
        \label{fig:example-preceq-revised}
    \end{figure}

    \note{
        \begin{itemize}
            \item figure 7
            \item maybe put the figure 5 from paper (original revision of <=) before this
            \item maybe put slides for complete example in here going from tpo to revision once, to revision twice
        \end{itemize}
    }
\end{frame}

\begin{frame}[fragile]{Courtroom Example: Conditional beliefs 1}
    \begin{figure}[H]
        \centering
        \begin{tikzpicture}[scale=1]
            \begin{axis}[
                    xticklabels={},
                    yticklabels={},
                    extra y ticks={1, 2, 3, 4, 5, 6, 7, 8},
                    extra y tick labels={$111$, $001$, $101$, $011$, $110$, $000$, $100$, $010$},
                    ytick style={draw=none},
                    xtick style={draw=none},
                    axis line style={draw=none}
                ]
                \addplot[
                    scatter,
                    scatter src=explicit symbolic,
                    mark size=3,
                    scatter/classes={
                            empty={mark=*, fill=white},
                            old={mark=x, fill=white}
                        },
                    nodes near coords*={\Label},
                    visualization depends on={value \thisrow{label} \as \Label}
                ]
                table [meta=class] {
                        x y class label

                        1 8 old \empty

                        3 8 empty
                        5 8 empty

                        1 7 old

                        3 7 empty
                        5 7 empty

                        2 6 old

                        4 6 empty
                        6 6 empty

                        2 5 old

                        4 5 empty
                        6 5 empty

                        5 4 empty
                        7 4 empty

                        5 3 empty
                        7 3 empty

                        6 2 empty
                        8 2 empty

                        6 1 empty
                        8 1 empty
                    };
            \end{axis}
        \end{tikzpicture}
        \caption{$\preceq_{p \ast \alpha}$ for $\alpha = r$}
        \label{fig:example-preceq-revised-alien}
    \end{figure}

    \note{
        \begin{itemize}
            \item figure 8
        \end{itemize}
    }
\end{frame}

\begin{frame}[fragile]{Courtroom Example: Conditional beliefs 2}
    \begin{figure}[H]
        \centering
        \begin{tikzpicture}[scale=1]
            \begin{axis}[
                    xticklabels={},
                    yticklabels={},
                    extra y ticks={1, 2, 3, 4, 5, 6, 7, 8},
                    extra y tick labels={$111$, $001$, $101$, $011$, $110$, $000$, $100$, $010$},
                    ytick style={draw=none},
                    xtick style={draw=none},
                    axis line style={draw=none}
                ]
                \addplot[
                    scatter,
                    scatter src=explicit symbolic,
                    mark size=3,
                    scatter/classes={
                            empty={mark=*, fill=white},
                            filled={mark=*, fill=black},
                            old={mark=x, fill=white}
                        },
                    nodes near coords*={\Label},
                    visualization depends on={value \thisrow{label} \as \Label}
                ]
                table [meta=class] {
                        x y class label

                        1 8 old \empty

                        3 8 filled
                        5 8 empty

                        1 7 old

                        3 7 filled
                        5 7 empty

                        2 6 old

                        4 6 empty
                        6 6 filled

                        2 5 old

                        4 5 empty
                        6 5 filled

                        5 4 filled
                        7 4 empty

                        5 3 filled
                        7 3 empty

                        6 2 empty
                        8 2 filled

                        6 1 empty
                        8 1 filled
                    };
            \end{axis}
        \end{tikzpicture}
        \caption{$\preceq_{p \ast \alpha \ast \beta}$ for $\beta = (p \vee q) \wedge (\neg p \vee \neg q)$}
        \label{fig:example-preceq-revised-alien2}
    \end{figure}

    \note{
        \begin{itemize}
            \item figure 9
        \end{itemize}
    }
\end{frame}

\begin{frame}{}
    \centering \Huge
    Thank You
\end{frame}
\note{
    \begin{itemize}
        \item Vorher: Revision der neuen Struktur $\preceq$, ein möglicher konkreter Beispieloperator + Visualisierung
        \item Jetzt: Fragen? Diskussion
    \end{itemize}
}

\begin{frame}[allowframebreaks=0.9]{References}
    \typeout{}
    \bibliographystyle{apalike}
    \bibliography{references}
\end{frame}

\end{document}